Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-03-23 PAPER/main.tex
added pros and cons to heuristics

\subsection{FFD Approx}
\label{sec:ffdapprox}
The first heuristic to be discussed is the Score-Constrained First-Fit Decreasing (SCFFD) heuristic, which is an extension of the original FFD. The items in $\mathcal{I}$ are, as usual, sorted in non-increasing order of widths, with ties broken by selecting the item with the smallest score width. Then, for each item $i \in \mathcal{I}$, SCFFD finds the lowest-indexed strip $S_j$ that can accommodate $i$ without being overfilled. If this strip $S_j$ is empty, place $i$ onto $S_j$ in a regular orientation, and add $S_j$ to the solution set $\mathcal{S}$. However, if $S_j$ is not empty, i.e. $S_j \in \mathcal{S}$, the algorithm checks if the vicinal sum constraint is met between the right-most score width on $S_j$ and one of the two score widths on $i$, assessing the smaller score width $a_i$ first. In either case, if the constraint is met, $i$ is placed accordingly onto $S_j$. Indeed, an item $i$ may not fulfil the vicinal sum constraint in any orientation, in which case SCFFD finds the next lowest-indexed strip and attempts to pack $i$ again. This process continues until every item $i \in \mathcal{I}$ has been placed onto a strip $S_j \in \mathcal{S}$.

\subsection{Smallest First}
\label{sec:ffdsmall}
This next heuristic focusses on packing each strip in turn, rather than each item as in the previous heuristic. Each new strip $S_j$ is initialised by placing the item $i \in \mathcal{I}'$ with the smallest score width onto $S_j$ in a regular position, where $\mathcal{I}'$ is the set of items that have not yet been packed onto a strip. The heuristic then continues to pack items onto $S_j$, by choosing an item in $\mathcal{I}'$ with the smallest score width that fulfils the vicinal sum constraint when placed alongside the right-most score width on $S_j$, and that does not exceed the capacity of $S_j$. Once there are no more items in $\mathcal{I}'$ that can be feasibly packed onto $S_j$, the heuristic adds $S_j$ to the solution set $\mathcal{S}$, and the next new strip $S_{j+1}$ is initialised. When $|\mathcal{I}'| = 0$, all items have been packed onto a strip, and the process terminates.

\subsection{FFD Exact}
\label{sec:ffdexact}
The last heuristic in this paper is formed by integrating the Alternating Hamiltonian Construction Algorithm (AHCA) detailed in Section ref{} with the classical FFD heuristic. Again, all items in $\mathcal{I}$ are sorted into non-increasing order of widths, with ties broken by choosing the item with the smallest score width. For each item $i \in \mathcal{I}$, the heuristic finds the lowest-indexed feasible strip $S_j$. As in the first heuristic, if $S_j$ is empty, $i$ is placed in a regular orientation onto $S_j$, and $S_j$ is then added to the solution set $\mathcal{S}$. Unlike the first heuristic, however, if $S_j$ is not empty ($S_j \in \mathcal{S}$), AHCA is applied on all items in $S_j$ and the current item $i$. If AHCA is successful, the items currently on $S_j$ are replaced by the new arrangement, which of course includes $i$. In the event that AHCA is unable to produce a solution, $S_j$ remains intact, and the heuristic attempts to pack $i$ onto the next lowest-indexed feasible strip.

\begin{figure}[h!]	
	\centering
	\includestandalone[width=\textwidth]{figures/strips}
	\caption{\textcolor{red}{BasicFFD vs. FFDincAHCA}}	
	\label{fig:ffdvsffdincahca}
\end{figure}

>>>>

\subsection{FFD Approx}
\label{sec:ffdapprox}
The first heuristic to be discussed is the Score-Constrained First-Fit Decreasing (SCFFD) heuristic, which is an extension of the original FFD. The items in $\mathcal{I}$ are, as usual, sorted in non-increasing order of widths, with ties broken by selecting the item with the smallest score width. Then, for each item $i \in \mathcal{I}$, SCFFD finds the lowest-indexed strip $S_j$ that can accommodate $i$ without being overfilled. If this strip $S_j$ is empty, place $i$ onto $S_j$ in a regular orientation, and add $S_j$ to the solution set $\mathcal{S}$. However, if $S_j$ is not empty, i.e. $S_j \in \mathcal{S}$, the algorithm checks if the vicinal sum constraint is met between the right-most score width on $S_j$ and one of the two score widths on $i$, assessing the smaller score width $a_i$ first. In either case, if the constraint is met, $i$ is placed accordingly onto $S_j$. Indeed, an item $i$ may not fulfil the vicinal sum constraint in any orientation, in which case SCFFD finds the next lowest-indexed strip and attempts to pack $i$ again. This process continues until every item $i \in \mathcal{I}$ has been placed onto a strip $S_j \in \mathcal{S}$. The additional check of the vicinal sum constraints and rotation of the items can be performed in constant time, and as such this heuristic has the same time complexity as FFD, $O(n \log n)$. This heuristic only checks the current items score widths against the right-most score width on a strip, and if the vicinal sum constraint is not met, the item cannot be placed on the strip. However, there may be another location on the strip that the item could be placed in feasibly, or a rearrangement of all items on the strip that produce a feasible alignment.

\subsection{Smallest First}
\label{sec:ffdsmall}
This next heuristic focuses on packing each strip in turn, rather than each item as in the previous heuristic. Each new strip $S_j$ is initialised by placing the item $i \in \mathcal{I}'$ with the smallest score width onto $S_j$ in a regular position, where $\mathcal{I}'$ is the set of items that have not yet been packed onto a strip. The heuristic then continues to pack items onto $S_j$, by choosing an item in $\mathcal{I}'$ with the smallest score width that fulfils the vicinal sum constraint when placed alongside the right-most score width on $S_j$, and that does not exceed the capacity of $S_j$. Once there are no more items in $\mathcal{I}'$ that can be feasibly packed onto $S_j$, the heuristic adds $S_j$ to the solution set $\mathcal{S}$, and the next new strip $S_{j+1}$ is initialised. When $|\mathcal{I}'| = 0$, all items have been packed onto a strip, and the process terminates. This heuristic is designed to align score widths such that their sum is as close to the minimum scoring distance as possible, which eliminates the possibility of aligning larger score widths unecessarily. The larger score widths are then available to be aligned next to smaller score widths. Heuristic prioritises vicinal sum constraint over item widths.

\subsection{FFD Exact}
\label{sec:ffdexact}
The last heuristic in this paper is formed by integrating the Alternating Hamiltonian Construction Algorithm (AHCA) detailed in Section ref{} with the classical FFD heuristic. Again, all items in $\mathcal{I}$ are sorted into non-increasing order of widths, with ties broken by choosing the item with the smallest score width. For each item $i \in \mathcal{I}$, the heuristic finds the lowest-indexed feasible strip $S_j$. As in the first heuristic, if $S_j$ is empty, $i$ is placed in a regular orientation onto $S_j$, and $S_j$ is then added to the solution set $\mathcal{S}$. Unlike the first heuristic, however, if $S_j$ is not empty ($S_j \in \mathcal{S}$), AHCA is applied on all items in $S_j$ and the current item $i$. If AHCA is successful, the items currently on $S_j$ are replaced by the new arrangement, which of course includes $i$. In the event that AHCA is unable to produce a solution, $S_j$ remains intact, and the heuristic attempts to pack $i$ onto the next lowest-indexed feasible strip. Due to the inclusion of AHCA, this heuristic is slower than the previous heuristics, with complexity $O(n^2)$. This heuristic has a significant advantage, however, as it is able to rearrange items entirely in order to find a feasible solution. In addition, we know that if a feasible alignment for a given set of items exists, AHCA is guaranteed to find and produce this solution. Thus, every subproblem in this heuristic, i.e. every time an item needs to be placed on a strip, can be solved exactly. For example, Figure \ref{fig:comparestrips} illustrates an instance of the SCSPP in which an item can feasibly fit onto a strip, i.e. the addition of the item onto the strip does not cause the strip's capacity to be exceed. In the original version of FFD, the items on the strip are, as expected, in non-increasing order, and the right-most score width has a value of 3. Regardless of orientation, the current item cannot meet the vicinal sum constraint. The heuristic will therefore attempt to place the item onto another strip, or, if there are no other open strips, a new strip will be required to accommodate the item, thus increasing the total number of strips. However, by using FFDincAHCA, a feasible alignment of the items including the current item exists, and hence the item does not need to be placed elsewhere, minimising the risk of potentially adding a new strip to the solution.

\begin{figure}[h!]	
	\centering
	\includestandalone[width=\textwidth]{figures/strips}
	\caption{Comparison of two heuristics, maximum strip length is 18, $\tau = 70$.}	
	\label{fig:comparestrips}
\end{figure}

\section{Comparison}
In order to provide a complete overview of the heuristics, sets of 100, 500, and 1000 items were used on strips of lengths 1500, 2500, and 5000. Item widths were randomly selected between 150 and 1000, and score widths between 1 and 70. This ensures that each item has two score lines.
